% This file contains acknowledgements
\vspace*{7cm}
\begin{center}
\textit{To those who dream of making the impossible possible}
\end{center}

\newpage

The work and contributions highlighted in this thesis are the result of
fruitful collaboration with a diversity of people. People who provided me
with advice, with motivation, with competition and enthusiasm. I have to thank
everyone who stood by me for the past years and supported my research and my
professional activity and, at the same time, tolerated my sometimes abrassive
behavior and harsh attitude.

Firstly, I am in great debt to the effort and careful pieces of advice
provided by my supervisor, prof. Nicolae Țăpuș. His calm attitude and
insightful remarks have been of exellent value; I owe him a lot of gratitude
for the directions he pointed out and the support he offered. Our
research collaboration was strenghten by working together in the P2P-Next
project where most of my research work had taken place.

Huge thanks go to my friend, colleague and collaborator, Răzvan Rughiniș.
Răzvan has been on my side for the whole duration of my PhD and has provided
me with constant motivation to push forward. Though we've had our share of
differences, Răzvan has never failed to provide me his full support towards
doing good quality research and putting it to good use.

My ``PhD brethren'', Mircea Bardac and George Milescu have been my major
collaborators and fellow researchers in the Peer-to-Peer field. All three of
us have worked closely and collaborated on many research tracks and published
many scientific articles together. We have learned a lot from each other. I
thank Mircea for providing us with new information regarding research tools
and applications. I thank George for his neverending effort and perfectionism
and his constant desire to do things the right way.

Working within the P2P-Next project I've had the pleasure of working close to
high quality researchers and engineers. People such as Johan Pouwelse, Keith
Mitchell, Johnathan Ishmael have been instrumental in my learning of clever
approaches to researching and skills required to be part of a large project.

With a high level of expertise in distributed systems research, prof. Valentin
Cristea provided me with valuable pieces of advice and directions for my PhD.
Prof. Cristea managed to transfer me some of his passion for research
activities and his insight on dealing with scientific challenges.

Andrei Pitiș and Tavi Purdilă have ensured mentorship in team-based
environments and technical matters. Their vast knowledge and skills and their
neverending trust in me mean a lot and have acted as a continuous boost of
enthusiasm.

Within the P2P-Next project in particular and the Peer-tp-Peer research area,
collaborators such as Călin Burloiu, Adriana Drăghici, Marius Sandu-Popa,
Costin Lupu, Aurelian Bogdan, Bogdan Druțu, Oana Baron supported my research
and allowed the birth of new ideas and expansion of old ones.

Thanks go to my colleagues and fellow researchers Laura Gheorghe, Mugurel
Andreica, Alex Costan, Eliana Tîrșa, Vlad Posea, Traian Rebedea, Andreea
Urzică. We've interacted plenty of times in various projects and exchanged
ideas, opinions and skills.

As part of the Systems group, people such as Vlad Dogaru, Alex Juncu, Daniel
Rosner, Laura Gheorghe, Mihai Bucicoiu, Bogdan Doinea have taken part in
fruitful discussions regarding research tracks. The diversity and activism of
the group are a constant source of motivation for my taking on as many
challenges as possible.

Working close to clever and dilligent students in engineering and research
projects has been instrumental to my constant yearn to try out new things.
Many thanks go to people such as Andrei Buhaiu, Lucian Cojocar, Alex Eftimie,
Vali Priescu, Cătălin Moraru, Sergiu Iordache. We've worked together on a
variety of projects that are continuing today with more and more people and
ideas.

I thank all members of the ROSEdu community for the constant involvement in
open-source and educational projects, for taking up challenges and for helping
me whenever possible. I thank Mihai Maruseac, Andrada Georgescu, Laura
Vasilescu, Victor Cărbune, Adrian Scoică, Andrei Maruseac, Vlad Voicu, Dragoș
Dena and others for supporting me in all good or bad ideas I've come up with
and providing the best outcome to adjiacent activities.

My family has and played a big part in my life and professional activity.
Their constant support, unending trust and continuous possitive attitude form
an infinite reservoir of energy for all activities I'm involved in. I thank my
parents, Ion and Cornelia, my sister Alina and my brother-in-law and good
friend Alex for always being there for me and their providing of help whenever
required.

For the past years, I've had the privilege to meet, interact and collaborate
with a diversity of people. I've learned a lot, gathered experience and
received constat support from the ever growing crowd of enthusiastic and
passionate people around. Thank you all for your patience, your stamina and
your trust in me. This work is embued with your effort and I will strive to
give you back my whole appreciation.
